PLAE enables quick and easy searching across one million ocular transcriptomes

### 2022-06-02 manuscript_00.Rmd
new diff tables

scEiaD_2020_v00 <- pool::dbPool(RSQLite::SQLite(), dbname =  glue('{data_dir_vGiga}/MOARTABLES__anthology_limmaFALSE___5000-transform-counts-universe-batch-scVIprojectionSO-8-0.1-50-5.sqlite'))
scEiaD_2020_v01 <- pool::dbPool(RSQLite::SQLite(), dbname =  glue('{data_dir_vPLAE}/MOARTABLES__anthology_limmaFALSE___4000-counts-universe-study_accession-scANVIprojection-15-5-0.1-50-20.sqlite'))

@@

num1 <- meta_filter %>% filter(Organ == 'Eye') %>% pull(Tissue) %>% unique() %>% length()
num2 <- meta_filter %>% filter(Organ != 'Eye') %>% pull(Tissue) %>% unique() %>% length()
nun3 <- meta_filter %>% filter(Organ == 'Eye') %>% pull(CellType) %>% unique() %>% length()
scEiaD_data_overview_cap <- fig_cap(name = 'scEiaD_data_overview', caption = glue::glue("Distribution of tissues and cell types in scEiaD. A. tabulation of the number of cells present across {num1} ocular and {num2} non-ocular tissues, B. Number of studies for each ocular tissue. C. Number of cells present across {num3} curated ocular-derived cell types."))

@@

## Differential testing identifies thousands genes frequently enriched across various ocular cell types 
```{r}
scEiaD_2020_v01 %>% tbl('wilcox_diff_testing') %>% 
  filter(Group == 'CellType (Predict)', FDR < 1e-5, mean_auc > 0.7 ) %>% 
  filter(Base %in% (meta_filter %>% filter(Organ == 'Eye') %>% pull(CellType_predict) %>% unique())) %>% 
  group_by(Base) %>% 
  summarise(Count = n()) %>% 
  rename(`CellType (Predict)` = Base, `Differentially Expressed Gene Count` = Count) %>% 
  as_tibble() %>% 
  flextable()
```

>>>>

scEiaD_2020_v00 <- pool::dbPool(RSQLite::SQLite(), dbname =  glue('{data_dir_vGiga}/MOARTABLES__anthology_limmaFALSE___5000-transform-counts-universe-batch-scVIprojectionSO-8-0.1-50-5.sqlite'))
scEiaD_2020_v01 <- pool::dbPool(RSQLite::SQLite(), dbname =  glue('{data_dir_vPLAE}/MOARTABLES__anthology_limmaFALSE___4000-counts-universe-study_accession-scANVIprojection-15-5-0.1-50-20__pointRelease01.sqlite'))

@@

num1 <- meta_filter %>% filter(Organ == 'Eye') %>% pull(Tissue) %>% unique() %>% length()
num2 <- meta_filter %>% filter(Organ != 'Eye') %>% pull(Tissue) %>% unique() %>% length()
num3 <- meta_filter %>% filter(Organ == 'Eye') %>% pull(CellType) %>% unique() %>% length()
scEiaD_data_overview_cap <- fig_cap(name = 'scEiaD_data_overview', caption = glue::glue("Distribution of tissues and cell types in scEiaD. A. tabulation of the number of cells present across {num1} ocular and {num2} non-ocular tissues, B. Number of studies for each ocular tissue. C. Number of cells present across {num3} curated ocular-derived cell types."))

@@

## Differential testing identifies thousands genes frequently enriched across various ocular cell types 
```{r}
diff_tab <- scEiaD_2020_v01 %>% tbl('diff_testing') %>% 
  filter(Group == 'CellType (Predict)', padj < 1e-5, (log2FoldChange) > 2) %>% 
  collect() %>% 
  filter(Against == 'All', Base %in% (meta_filter %>% filter(Organ == 'Eye') %>% pull(CellType_predict) %>% unique()))

diff_tab %>% 
  group_by(Base, Organism) %>% 
  summarise(Count = n()) %>% 
  rename(`CellType (Predict)` = Base, `Differentially Expressed Gene Count` = Count) %>% 
  pivot_wider(values_from = `Differentially Expressed Gene Count`, names_from = Organism) %>% replace(is.na(.),0) %>% 
  as_tibble() %>% 
  flextable()
```

```{r}
# top 100 genes in each organism/base category
diff_tab %>%

  group_by(Base, Organism) %>% 
  top_n(., 100, -padj) %>% 
  arrange(Base, Organism) %>% 
  relocate(Base, Organism) %>% 
  select(-Against, -Group)
```

```{r}
# consistently diff genes
consistent <- diff_tab %>% 
  group_by(Gene, Base)  %>% 
  summarise(Count = n()) %>% 
  filter(Count > 1) %>% 
  mutate(k = paste(Gene, Base, sep = '_'))

diff_tab %>% mutate(k = paste(Gene, Base, sep = '_')) %>% 
  filter(k %in% consistent$k) %>% 
  group_by(Base, Organism) %>% 
  summarise(Count = n()) %>% 
  rename(`CellType (Predict)` = Base, `Differentially Expressed Gene Count` = Count) %>% 
  pivot_wider(values_from = `Differentially Expressed Gene Count`, names_from = Organism) %>% replace(is.na(.),0) %>% 
  as_tibble() %>% 
  flextable()

```{r}
# flip flop diff genes
down_tab <- scEiaD_2020_v01 %>% tbl('diff_testing') %>% 
  filter(Group == 'CellType (Predict)', padj < 1e-5, (log2FoldChange) < -2) %>% 
  collect() %>% 
  filter(Against == 'All', Base %in% (meta_filter %>% filter(Organ == 'Eye') %>% pull(CellType_predict) %>% unique()))

diff_tab %>% mutate(k = paste(Gene, Base, sep = '_')) %>% 
  filter(k %in% consistent$k) %>% 
  group_by(Base, Organism) %>% 
  summarise(Count = n()) %>% 
  rename(`CellType (Predict)` = Base, `Differentially Expressed Gene Count` = Count) %>% 
  pivot_wider(values_from = `Differentially Expressed Gene Count`, names_from = Organism) %>% replace(is.na(.),0) %>% 
  as_tibble() %>% 
  flextable()
